---
title: "Example Report: DEMO OF COMMON ERRORS"
output: prettydoc::html_pretty
author: "GRAPH Courses team"
---

We'll use this document to show you some common errors and issues you might encounter when writing your report. The issues will be debugged in a live interactive session during class.
# Road Injuries of Motorcyclists in Colombia 

```{r setup}
# Load packages
if (!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, here, janitor, esquisse)
```

```{r import-data}
col_accidents <- read_csv(here("data/EXAMPLE_moto_accidents_colombia.csv"))
col_accidents_clean <- col_accidents %>% clean_names()
```

The dataset analysed in this report provides information about `r nrow(col_accidents_clean)` deaths caused by motorcycle accidents in Medellín, a Colombian city, from 2012 to 2015. 

## Age group distribution

```{r age-plot}
esquisser(col_accidents_clean)
age_group_plot <-
  col_accidents_clean %>%
  ggplot() +
  aes(x = age_group) +
  geom_bar(fill = "#678BCB") +
  labs(title = "Age group distribution of deaths")
```

```{r age-table}
freq_df <- tabyl(col_accidents_clean, age_group)
flex <- flextable(freq_df)
flex
```

