---
title: "Example Report: Motorcycle Accidents in Colombia"
output: prettydoc::html_pretty
author: "GRAPH Courses team"
---

This document is a model example of what a report for this assignment can look like. You may use this document as a guide on how to write your report.

# Road Injuries of Motorcyclists in Colombia 

```{r setup, echo = FALSE, message = FALSE}
# Load packages
if (!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, here, janitor, esquisse, flextable)
```

```{r import, echo = FALSE, message = FALSE}
# Import CSV
col_accidents <- read_csv(here("data/EXAMPLE_moto_accidents_colombia.csv"))
# Clean variable names
col_accidents_clean <- clean_names(col_accidents)
```

The dataset analysed in this report provides information about `r nrow(col_accidents_clean)` deaths caused by motorcycle accidents in Medellín, a Colombian city, from 2012 to 2015. The data is sourced from medical and police certificates and can be accessed [here](https://zenodo.org/record/4836304).

## Age group distribution

I chose to look at the age group distribution of fatalities from motorcycle accidents. The victims ranged from `r min(col_accidents_clean$age)` to `r max(col_accidents_clean$age)` years old.

```{r age-plot, echo = FALSE}
age_group_plot <-
  col_accidents_clean %>%
  ggplot() +
  aes(x = age_group) +
  geom_bar(fill = "#678BCB") +
  labs(title = "Age group distribution of deaths")
age_group_plot
```

The table below summarizes the data:

```{r age-table, echo = FALSE}
freq_df <- tabyl(col_accidents_clean, age_group)
flex <- flextable(freq_df)
flex_with_caption <- set_caption(flex, caption = "Age group distribution of deaths")
flex_with_caption
```

The 5-year age group with the most deaths was the 20 to 24 age group.
