---
title: "Workshop Exercise: Joining and Pivoting"
author: "YOUR NAME HERE" # Remember to add your name!
---

# Introduction

In this workshop, you'll learn how to clean and analyze HIV and AIDS data across different countries and years. We will start with raw datasets containing estimates of new and total HIV cases per country-year. The goal is to clean and merge these datasets with population data.

Please refer to the lesson notes to aid you in the workshop exercise:

- [Intro to Pivoting Data](https://the-graph-courses.github.io/data_untangled_wp/lessons/ls07_pivoting.html)
- [Intro to Joining Datasets](https://the-graph-courses.github.io/fdar/FDAR_EN_joining_1/FDAR_EN_joining_1.html)

# Load Libraries

```{r}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, here, countrycode, esquisse)
```

# Data Import

We will work with two raw datasets, each containing estimates per country-year. These data were accessed from the Gapminder foundation, at [www.gapminder.org/data](https://www.gapminder.org/data).

1. **People living with HIV:** This dataset contains the total number of people currently infected with HIV (*data/hiv_prevalence.csv*).
2. **New HIV infections:** This dataset provides the total number of people newly infected with HIV during a given year (*data/hiv_incidence.csv*).

Load the datasets into R using `read_csv()`.

```{r}
# Load raw data from specified paths
total_hiv_raw <- read_csv(here("data/hiv_prevalence.csv"))
new_hiv_raw <- read_csv(here("data/hiv_incidence.csv"))
```

# Explore the Datasets

Use functions like `head()`, `View()`, and `summary()` to inspect the datasets.

```{r}
# Your code here
```

# Data Pivoting

Now that we have imported the data it's time to pivot it. 

The datasets are currently in wide format, with each year as a separate column, something like this:

| country     | 1990  | 1991  | 1992  | 1993  | 1994  |
|-------------|-------|-------|-------|-------|-------|
| Afghanistan | 600   | 600   | 1000  | 1100  | 1200  |
| Angola      | 33000 | 39000 | 46000 | 54000 | 62000 |
| Argentina   | 27000 | 31000 | 35000 | 39000 | 43000 |
| Armenia     | 350   | 600   | 600   | 600   | 1000  |
| Australia   | 12000 | 12000 | 12000 | 11000 | 11000 |
| Austria     | 600   | 600   | 1000  | 1300  | 1500  |

We need to convert this to long format, where each row represents a single year-country combination. For example:

| country     | year | cases |
|-------------|------|-------|
| Afghanistan | 1990 | 600   |
| Afghanistan | 1991 | 600   |
| Afghanistan | 1992 | 1000  |
| Afghanistan | 1993 | 1100  |
| Afghanistan | 1994 | 1200  |

## Instructions for Pivoting `new_hiv_raw` Data

Reshape the `new_hiv_raw` dataset by converting the columns that represent different years into a new column called `year`, and place the corresponding values into a new column called `new_cases`.

Hints:

1. Use `pivot_longer()` on `new_hiv_raw`.
2. Exclude the `country` and `country_code` columns from being pivoted by specifying `cols = -c(country, country_code)`.
3. Set the names of the new columns inside `pivot_longer()` using:
   - `names_to = "year"` to label the column that will store the years.
   - `values_to = "new_cases"` to store the values representing new HIV cases under the `new_cases` column.

```{r}
# Your code here
new_hiv_long <- 
```

---

## Instructions for Pivoting `total_hiv_raw` Data

In a similar manner, pivot the `total_hiv_raw` dataset to a long format. Convert the year columns into a new column called `year` and the counts into `total_cases`.

```{r}
# Your code here
total_hiv_long <- 
```

**CHECKPOINT:**  
After performing these steps, make sure that you have two reshaped datasets:
- `new_hiv_long` with columns: `country`, `country_code`, `year`, and `new_cases`.
- `total_hiv_long` with columns: `country`, `country_code`, `year`, and `total_cases`.

# Joining the HIV Datasets

Now that both datasets are in long format, we can join them. Note that you need to use the `country`, `country_code`, and `year` columns as the keys (`by = c("country", "country_code", "year")`).

Use `full_join()` so that we can keep as much information as possible.

```{r}
# Your code here
hiv_data_combined <- 
```

**CHECKPOINT:** The combined dataset should have columns for `country`, `country_code`, `year`, `total_cases`, and `new_cases`.

# Add Population Data

The `world_bank_pop` dataset from **tidyr** contains information on population per year from 2000 to 2017.

```{r}
tidyr::world_bank_pop # Run this code to see the dataset
```

## Filter and Pivot Population Data

Filter to just the `SP.POP.TOTL` indicator, then pivot the data to a long format. Store the result as `population_long`.

Rename the "country" column to "country_code" so that we can join it properly later on. 

```{r}
# Your code here
```

## Join Population Data with HIV Data

Use `left_join()` to add the population information to your HIV data combined dataset (the HIV data should be the left or primary dataset). You should use both "country_code" and "year" as keys.

```{r}
# Your code here
hiv_data_combined_pop <- 
```

**CHECKPOINT:** Verify that `hiv_data_combined_pop` now contains a `population` column.

## Calculate Incidence Rate and Prevalence Rate

Now, calculate the incidence rate and prevalence rate:

- **Incidence Rate** = (New Cases / Population) * 100,000
- **Prevalence Rate** = (Total Cases / Population) * 100,000

```{r}
# Your code here
```

# Plot Incidence Rates

Now, you'll select 3 countries of your choice and plot their incidence rate over time. 

Before passing your data into esquisse, you should

- filter the data to narrow down the countries. 
- convert the `year` column to a `numeric` type with `mutate()` and `as.numeric()`.

```{r}
# Your code here
```

# Plot Prevalence Rates

Again, select 3 countries of your choice and plot their prevalence rate over time. 

```{r}
# Your code here
```

# Optional Ungraded Challenge

In this challenge, you will:

1. **Add Regional Information:** Use the `countrycode` package to get regional (continent) information for each country.
2. **Merge Continent Data:** Merge the continent data with your dataset.
3. **Identify Highest Prevalence Rates per Continent:** Find the countries with the highest prevalence rate in 2011 for each continent.
4. **Create a Bar Chart:** Make a bar chart comparing the 2011 prevalence rates of these countries.

```{r}
# Your code here
countrycode::codelist %>%
 select(country.name.en, continent)
```
